Generation of TypeScript Declaration Files from JavaScript Code

### 2020-07-30 programming-2021/main.tex
intro

\section{Introduction}
\label{sec:introduction}
JavaScript has become the most popular language for writing web
applications \cite{github-statistics}. It is also gaining popularity
for back-end applications running in \NodeJS{}, a JavaScript-based
server-side platform. JavaScript is appealing to developers because
its forgiving dynamic typing enables 
them to create simple pieces of code very quickly and proceed on a
trial-and-error basis.

JavaScript was never intended to be more than a
scripting language. Hence, it lacks features for maintaining and evolving large
codebases.
Presently, however, JavaScript is being used for creating large and complex
applications. 
Mistakes such as mistyped property
names and misunderstood or unexpected type coercions cause developers
to spend a significant amount of time in debugging. There is ample
evidence for such mishaps. For example, a 
JavaScript code blog\footnote{\url{https://wtfjs.com}} collects experiences
from developers facing unexpected situations while programming in
JavaScript. \coderef{code:introduction-javascript-wtfs} exposes some
of these unintuitive JavaScript behaviors.

@@

The cognitive load produced by such dynamic typing (which includes
unchecked property names) is not present in other languages that use
build tools based on type information. This insight motivated the
creation of TypeScript, a superset of JavaScript with type
annotations \cite{typescript}. It has become a widely used alternative
among JavaScript developers, because it incorporates features that are
helpful for developing and maintaining large applications
\cite{DBLP:conf/icse/GaoBB17}. TypeScript enables the early detection
of several kinds of run-time errors and the integration of code intelligence
tools like auto-completion in an IDE.

Existing JavaScript libraries can be used in a TypeScript project by
adding a declaration file that contains a description of the library's
API in terms of types. Lots of declaration files for popular
JavaScript libraries have been created in a community effort in the
DefinitelyTyped repository~\cite{definitely-typed-repository}.
At the time of writing this repository contains declaration files for
more than 6000 JavaScript libraries. Unfortunately, most declaration
files in this libraries have been manually created and maintained,
which is error prone and time consuming. As TypeScript takes a
declaration file at face value, discrepancies are not detected at
compile time and the inaccurate code-intelligence features are misleading the
programmer. Moreover, TypeScript does not perform any run-time 
checking on types, either, so that a discrepancy  between the declaration file
and its corresponding JavaScript library can lead to unexpected 
behavior and crashes. Such behavior can lead to developer
frustration, longish debugging sessions, and decreasing confidence in
the tool chain.

Some previous work tackled the problem of improving the quality of
declaration files. Feldthaus et al
\cite{DBLP:conf/oopsla/FeldthausM14} search automatically for
mismatches between a declaration file and implementation code. TSTest
\cite{DBLP:journals/pacmpl/KristensenM17}
adapted feedback-directed random testing to detect discrepancies between a
declaration file and a JavaScript library. Tools like TSInfer and
TSEvolve~\cite{DBLP:conf/fase/KristensenM17} are designed for
assisting the creation of new declaration 
files and supporting the evolution the declaration file when the
corresponding JavaScript library gets modified, respectively. They
rely on an existing static analyzer for JavaScript \cite{DBLP:conf/sas/JensenMT09}.
TypeScript itself
developed \texttt{dts-gen}, a tool that generates a
declaration file that is meant to be used as a \emph{starting point
for writing a high-quality declaration file} \cite{dts-gen}.

In this work we present the tool \texttt{dts-generate}
as a first step to explore the possibilities for
generating useful declaration files without the heavy lifting of
static analysis. \texttt{dts-generate} comes with a framework that
supports the generation of declaration files for an existing
JavaScript library published to the \NPM{} registry. The tool gathers
data flow and type information at run-time to generate a declaration
file based on that information.

>>>>

\section{Introduction}
\label{sec:introduction}
JavaScript is the most popular language for writing web
applications \cite{github-statistics}. It is also gaining popularity
for back-end applications running in \NodeJS{}, a JavaScript-based
server-side platform. JavaScript is appealing to developers because
its forgiving dynamic typing enables 
them to create simple pieces of code very quickly and proceed on a
trial-and-error basis.

JavaScript was never intended to be more than a
scripting language and thus lacks features for maintaining and evolving large
codebases. However, JavaScript is increasingly used for creating large and complex
applications. 
Mistakes such as mistyped property
names and misunderstood or unexpected type coercions cause developers
to spend a significant amount of time in debugging. There is ample
evidence for such mishaps. For example, a 
JavaScript code blog\footnote{\url{https://wtfjs.com}} collects experiences
from developers facing unexpected situations while programming in
JavaScript. \coderef{code:introduction-javascript-wtfs} exposes some
of these unintuitive JavaScript behaviors.

@@

The cognitive load produced by these behaviors is mitigated in languages that use
build tools based on type information. This insight motivated the
creation of TypeScript, a superset of JavaScript with expressive type
annotations \cite{typescript}. It has become a widely used alternative
among JavaScript developers, because it incorporates features that are
helpful for developing and maintaining large applications
\cite{DBLP:conf/icse/GaoBB17}. TypeScript enables the early detection
of several kinds of run-time errors and the integration of code intelligence
tooling like auto-completion in an IDE.

Despite the advantages, it is unrealistic to expect the world to
switch to TypeScript in a day. Therefore, 
existing JavaScript libraries can be used in a TypeScript project by
adding a declaration file that contains a description of the library's
API in terms of types. Lots of declaration files for popular
JavaScript libraries have been created in a community effort in the
DefinitelyTyped repository~\cite{definitely-typed-repository}.
At the time of writing this repository contains declaration files for
more than 6000 JavaScript libraries. Unfortunately, most declaration
files in this repository have been manually created and maintained,
which is error prone and time consuming. As TypeScript takes a
declaration file at face value, discrepancies are not detected at
compile time and the resulting inaccurate code-intelligence features
are confusing for the
programmer. Moreover, TypeScript does not perform any run-time 
checking of types, either, so that a discrepancy  between the declaration file
and its corresponding JavaScript library can lead to unexpected 
behavior and crashes. Such behavior can lead to developer
frustration, longish debugging sessions, and decreasing confidence in
the tool chain.

We aim to improve on this situation by automatically generating
sound TypeScript declaration files. To this end, we rely on the examples provided
by programmers as part of the documentation of their library. If these
examples are sufficiently rich, then our toolchain can generate the
desired declaration files from them. This approach takes advantage of
best practices in the dynamic languages community which favors
examples over writing out type signatues.

\begin{description}
\item[PJT: How about a short example at this point?] 
\item[PJT: maybe later] Some previous work tackled the problem of improving
  the quality of declaration files. Feldthaus et al
  \cite{DBLP:conf/oopsla/FeldthausM14} search automatically for
  mismatches between a declaration file and implementation
  code. TSTest \cite{DBLP:journals/pacmpl/KristensenM17} adapted
  feedback-directed random testing to detect discrepancies between a
  declaration file and a JavaScript library. Tools like TSInfer and
  TSEvolve~\cite{DBLP:conf/fase/KristensenM17} are designed for
  assisting the creation of new declaration files and supporting the
  evolution the declaration file when the corresponding JavaScript
  library gets modified, respectively. They rely on an existing static
  analyzer for JavaScript \cite{DBLP:conf/sas/JensenMT09}.  TypeScript
  itself developed \texttt{dts-gen}, a tool that generates a
  declaration file that is meant to be used as a \emph{starting point
    for writing a high-quality declaration file} \cite{dts-gen}.
\end{description}
In this work we present the tool \texttt{dts-generate}
as a first step to explore the possibilities for
generating useful declaration files without the heavy lifting of
static analysis. \texttt{dts-generate} comes with a framework that
supports the generation of declaration files for an existing
JavaScript library published to the \NPM{} registry. The tool gathers
data flow and type information at run-time to generate a declaration
file based on that information.
